---
title: "The Competitive Esports Physiological, Affective, and Video (CEPAV) Dataset"
author: "Maciej Behnke"
date: "2024-03-28"
output:
   html_document:  # Or another output format like pdf_document, word_document, etc.
    toc: true
    toc_float: true  # Optional: makes the TOC floating (html_document only)
    toc_depth: 3     # Optional: specify the depth of headers to include in the TOC
---

This is code for and the results of analysis included in the manuscript:
The Competitive Esports Physiological, Affective, and Video (CEPAV) Dataset

Link to OSF: https://osf.io/kgdsx/

Authors
Maciej Behnke1*, Wadim Krzyżaniak1, Jan Nowak2, Szymon Kupiński2, Patrycja Chwiłkowska1, Szymon Jęśko Białek1, Maciej Kłoskowski1, Patryk Maciejewski1, Kacper Szymański1,  Daniël Lakens3, Kate Petrova4, Jeremy P. Jamieson5, James J. Gross4


Affiliations
1 Faculty of Psychology and Cognitive Science, Adam Mickiewicz University, Poznań, Poland

2 Network Services Division, Poznan Supercomputing and Networking Center, Poznan, 61-139, Poland

3 Human-Technology Interaction, Eindhoven University of Technology, Eindhoven, The Netherlands

4 Department of Psychology, Stanford University, Stanford, USA

5 Department of Psychology, University of Rochester, Rochester, USA

corresponding author: Maciej Behnke (macbeh@amu.edu.pl)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, max.print = Inf)

options(repos = c(CRAN = "https://cran.r-project.org"))
options(pillar.print_max = 30)
```

# 1. Loading and installing all needed packages


```{r}


### remove all objects from workspace

rm(list=ls())

### load packages 
install.packages(c("tidyverse" 
,"readxl"
,"dplyr"
,"tibble"
,"readxl"
,"plyr"
, "careless"
, "psych"
, "data.table"
, "rempsyc"
, "lme4"
,"Matrix",
"papaja",
"afex",
"lavaan"))

library("tidyverse")
library("dplyr")
library('tibble')
library('readxl')
library("plyr")
library("careless")
library("psych")
library("data.table")
library("rempsyc")
library("lme4")
library("Matrix")
library("papaja")
library("afex")
library("lavaan")

```
# 2. Loading data 

```{r}

### set working directory
setwd("C:/Users/macbe/OneDrive/Behnke Dropbox/Upcoming Gaming Project/OPUS 20 Stypedyści/Sci Data/ms")

#import and store data from excel file
SMI_data2 <- read_excel("C:/Users/macbe/OneDrive/Behnke Dropbox/Upcoming Gaming Project/OPUS 20 Stypedyści/Sci Data/ms/CEPAV_data.xlsx", sheet = "self_reports")
CEPAV_data <- read_excel("C:/Users/macbe/OneDrive/Behnke Dropbox/Upcoming Gaming Project/OPUS 20 Stypedyści/Sci Data/ms/CEPAV_data.xlsx", sheet = "physio_behav")
SNR_data <- read_excel("C:/Users/macbe/OneDrive/Behnke Dropbox/Upcoming Gaming Project/OPUS 20 Stypedyści/Sci Data/ms/SNR_working.xlsx")

#View(SMI_data2)
#View(SNR_data)

```

# 3. Identifing low signal quality based on SNR based on MAD > 3

```{r}


SNR_data2 <- SNR_data %>% select(SNR_S1_ECG:SNR_S1_tr, SNR_S2_ECG:SNR_S2_tr)
outliers_SNR <- find_mad(SNR_data2, col.list = names(SNR_data2),  criteria = 3)
outliers_SNR

#View(SNR_data2)
#416 outling signals out of 10545 

```
# 4. Calculating mean SNRs for each collumn

```{r}
mean_values <- colMeans(SNR_data2, na.rm = TRUE)

# Print the mean values
print(mean_values)
print(min(mean_values))
print(max(mean_values))

min_value <- min(SNR_data2, na.rm = TRUE)
print(min_value)



```


# 5. Plotting avarages across Stage 1 & 3.

```{r}

HR_desc  <- CEPAV_data %>% select(baseline_visit1_min1_hr:training3_revery_min2_hr, baseline_visit2_min1_hr:tournament8_recovery_min2_hr)
HRV_desc  <- CEPAV_data %>% select(baseline_visit1_min1_hrv:training3_revery_min2_hrv, baseline_visit2_min1_hrv:tournament8_recovery_min2_hrv)
SBP_desc  <- CEPAV_data %>% select(baseline_visit1_min1_SBP:training3_revery_min2_SBP, baseline_visit2_min1_SBP:tournament8_recovery_min2_SBP)
DBP_desc  <- CEPAV_data %>% select(baseline_visit1_min1_DBP:training3_revery_min2_DBP, baseline_visit2_min1_DBP:tournament8_recovery_min2_DBP)
CO_desc  <- CEPAV_data %>% select(baseline_visit1_min1_CO:training3_revery_min2_CO, baseline_visit2_min1_CO:tournament8_recovery_min2_CO)
TPR_desc  <- CEPAV_data %>% select(baseline_visit1_min1_TPR:training3_revery_min2_TPR, baseline_visit2_min1_TPR:tournament8_recovery_min2_TPR)
Acc_WR_desc  <- CEPAV_data %>% select(baseline_visit1_min1_wr:training3_revery_min2_wr, baseline_visit2_min1_wr:tournament8_recovery_min2_wr)
Acc_TL_desc  <- CEPAV_data %>% select(baseline_visit1_min1_tl:training3_revery_min2_tl, baseline_visit2_min1_tl:tournament8_recovery_min2_tl)
Acc_TR_desc  <- CEPAV_data %>% select(baseline_visit1_min1_tr:training3_revery_min2_tr, baseline_visit2_min1_tr:tournament8_recovery_min2_tr)


Performance_desc  <- SMI_data2 %>% select(condition,
                                          Training1_Score,
                                          Training2_Score,
                                          Training3_Score,
                                         Tournament1_Score,
                                         Tournament2_Score,
                                         Tournament3_Score,
                                         Tournament4_Score,
                                         Tournament5_Score,
                                         Tournament6_Score,
                                         Tournament7_Score,
                                         Tournament8_Score
                                         )

Amus_desc  <- SMI_data2 %>% select(condition,
                                   Training1_Amusement,
                                          Training2_Amusement,
                                          Training3_Amusement,
                                     Tournament1_Amusement_T2,
                                  Tournament2_Amusement_T2,
                                  Tournament3_Amusement_T2,
                                  Tournament4_Amusement_T2,
                                  Tournament5_Amusement_T2,
                                  Tournament6_Amusement_T2,
                                  Tournament7_Amusement_T2,
                                  Tournament8_Amusement_T2
)

Joy_desc  <- SMI_data2 %>% select(condition,
                                  Training1_Joy,
                                          Training2_Joy,
                                          Training3_Joy,
                                  Tournament1_Joy_T2,
                                  Tournament2_Joy_T2,
                                  Tournament3_Joy_T2,
                                  Tournament4_Joy_T2,
                                  Tournament5_Joy_T2,
                                  Tournament6_Joy_T2,
                                  Tournament7_Joy_T2,
                                  Tournament8_Joy_T2
)

Exc_desc  <- SMI_data2 %>% select(condition,
                                  Training1_Excitement,
                                          Training2_Excitement,
                                          Training3_Excitement,
                                  Tournament1_Excitement_T2,
                                  Tournament2_Excitement_T2,
                                  Tournament3_Excitement_T2,
                                  Tournament4_Excitement_T2,
                                  Tournament5_Excitement_T2,
                                  Tournament6_Excitement_T2,
                                  Tournament7_Excitement_T2,
                                  Tournament8_Excitement_T2
)

Pride_desc  <- SMI_data2 %>% select(condition,
                                    Training1_Pride,
                                          Training2_Pride,
                                          Training3_Pride,
                                  Tournament1_Pride_T2,
                                  Tournament2_Pride_T2,
                                  Tournament3_Pride_T2,
                                  Tournament4_Pride_T2,
                                  Tournament5_Pride_T2,
                                  Tournament6_Pride_T2,
                                  Tournament7_Pride_T2,
                                  Tournament8_Pride_T2
)

Stress_desc  <- SMI_data2 %>% select(condition,
                                     Training1_Stress,
                                          Training2_Stress,
                                          Training3_Stress,
                                         Tournament1_Stress_T2,
                                         Tournament2_Stress_T2,
                                         Tournament3_Stress_T2,
                                         Tournament4_Stress_T2,
                                         Tournament5_Stress_T2,
                                         Tournament6_Stress_T2,
                                         Tournament7_Stress_T2,
                                         Tournament8_Stress_T2
)

Anger_desc  <- SMI_data2 %>% select(condition,
                                    Training1_Anger,
                                          Training2_Anger,
                                          Training3_Anger,
                                         Tournament1_Anger_T2,
                                         Tournament2_Anger_T2,
                                         Tournament3_Anger_T2,
                                         Tournament4_Anger_T2,
                                         Tournament5_Anger_T2,
                                         Tournament6_Anger_T2,
                                         Tournament7_Anger_T2,
                                         Tournament8_Anger_T2
)
Fear_desc  <- SMI_data2 %>% select(condition,
                                  Training1_Fear,
                                          Training2_Fear,
                                          Training3_Fear,
                                         Tournament1_Fear_T2,
                                         Tournament2_Fear_T2,
                                         Tournament3_Fear_T2,
                                         Tournament4_Fear_T2,
                                         Tournament5_Fear_T2,
                                         Tournament6_Fear_T2,
                                         Tournament7_Fear_T2,
                                         Tournament8_Fear_T2
)
Over_desc  <- SMI_data2 %>% select(condition,
                                   Training1_Overwhelm,
                                          Training2_Overwhelm,
                                          Training3_Overwhelm,
                                         Tournament1_Overwhelm_T2,
                                         Tournament2_Overwhelm_T2,
                                         Tournament3_Overwhelm_T2,
                                         Tournament4_Overwhelm_T2,
                                         Tournament5_Overwhelm_T2,
                                         Tournament6_Overwhelm_T2,
                                         Tournament7_Overwhelm_T2,
                                         Tournament8_Overwhelm_T2
)


HR_means <- sapply(HR_desc, function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)
HRV_means <- sapply(HRV_desc, function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)
SBP_means <- sapply(SBP_desc, function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)
DBP_means <- sapply(DBP_desc, function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)
CO_means <- sapply(CO_desc, function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)
TPR_means <- sapply(TPR_desc, function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)
Acc_WR_means <- sapply(Acc_WR_desc, function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)
Acc_TL_means <- sapply(Acc_TL_desc, function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)
Acc_TR_means <- sapply(Acc_TR_desc, function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)

Perf_means <- sapply(Performance_desc[-1], function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)


Amus_means <- sapply(Amus_desc[-1], function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)
Joy_means <- sapply(Joy_desc[-1], function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)
Exc_means <- sapply(Exc_desc[-1], function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)
Pride_means <- sapply(Pride_desc[-1], function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)
Stress_means <- sapply(Stress_desc[-1], function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)
Anger_means <- sapply(Anger_desc[-1], function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)
Fear_means <- sapply(Fear_desc[-1], function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)
Over_means <- sapply(Over_desc[-1], function(col) if(is.numeric(col)) mean(col, na.rm = TRUE) else NA)


# Define the custom x-axis labels and positions
custom_labels <- c(1, 12, 1, 12, 23, 34,45)
custom_positions <- c(1, 12, 24, 35, 47, 58,69)
highlight_points <- c(8, 9, 14, 15, 20, 21, 31, 32, 37, 38, 43, 44, 49, 50, 55, 56, 61, 62, 67, 68, 73, 74)

# HR Plot
plot(HR_means, main = "HR", xlab = "", ylab = "Mean Value", cex.main = 1.5, xaxt = "n")
lines(HR_means, type = "o", col = "#1f77b4", lwd = 2)
points(highlight_points, HR_means[highlight_points], col = "#ff7f7f", pch = 16, cex = 1.5)
abline(v = 23.5, lty = 2, col = "red", lwd = 2)
axis(1, at = custom_positions, labels = custom_labels)

# HRV Plot
plot(HRV_means, main = "HRV", xlab = "", ylab = "Mean Value", cex.main = 1.5, xaxt = "n")
lines(HRV_means, type = "o", col = "#1f77b4", lwd = 2)
points(highlight_points, HRV_means[highlight_points], col = "#ff7f7f", pch = 16, cex = 1.5)
abline(v = 23.5, lty = 2, col = "red", lwd = 2)
axis(1, at = custom_positions, labels = custom_labels)

# SBP Plot
plot(SBP_means, main = "SBP", xlab = "", ylab = "Mean Value", cex.main = 1.5, xaxt = "n")
lines(SBP_means, type = "o", col = "#1f77b4", lwd = 2)
points(highlight_points, SBP_means[highlight_points], col = "#ff7f7f", pch = 16, cex = 1.5)
abline(v = 23.5, lty = 2, col = "red", lwd = 2)
axis(1, at = custom_positions, labels = custom_labels)

# DBP Plot
plot(DBP_means, main = "DBP", xlab = "", ylab = "Mean Value", cex.main = 1.5, xaxt = "n")
lines(DBP_means, type = "o", col = "#1f77b4", lwd = 2)
points(highlight_points, DBP_means[highlight_points], col = "#ff7f7f", pch = 16, cex = 1.5)
abline(v = 23.5, lty = 2, col = "red", lwd = 2)
axis(1, at = custom_positions, labels = custom_labels)

# CO Plot
plot(CO_means, main = "CO", xlab = "", ylab = "Mean Value", cex.main = 1.5, xaxt = "n")
lines(CO_means, type = "o", col = "#1f77b4", lwd = 2)
points(highlight_points, CO_means[highlight_points], col = "#ff7f7f", pch = 16, cex = 1.5)
abline(v = 23.5, lty = 2, col = "red", lwd = 2)
axis(1, at = custom_positions, labels = custom_labels)

# TPR Plot
plot(TPR_means, main = "TPR", xlab = "", ylab = "Mean Value", cex.main = 1.5, xaxt = "n")
lines(TPR_means, type = "o", col = "#1f77b4", lwd = 2)
points(highlight_points, TPR_means[highlight_points], col = "#ff7f7f", pch = 16, cex = 1.5)
abline(v = 23.5, lty = 2, col = "red", lwd = 2)
axis(1, at = custom_positions, labels = custom_labels)

# Accelerometer Left Knee Plot
plot(Acc_TL_means, main = "Accelerometer Left Knee", xlab = "", ylab = "Mean Value", cex.main = 1.5, xaxt = "n")
lines(Acc_TL_means, type = "o", col = "#1f77b4", lwd = 2)
points(highlight_points, Acc_TL_means[highlight_points], col = "#ff7f7f", pch = 16, cex = 1.5)
abline(v = 23.5, lty = 2, col = "red", lwd = 2)
axis(1, at = custom_positions, labels = custom_labels)

# Accelerometer Right Knee Plot
plot(Acc_TR_means, main = "Accelerometer Right Knee", xlab = "", ylab = "Mean Value", cex.main = 1.5, xaxt = "n")
lines(Acc_TR_means, type = "o", col = "#1f77b4", lwd = 2)
points(highlight_points, Acc_TR_means[highlight_points], col = "#ff7f7f", pch = 16, cex = 1.5)
abline(v = 23.5, lty = 2, col = "red", lwd = 2)
axis(1, at = custom_positions, labels = custom_labels)

# Accelerometer Right Wrist Plot
plot(Acc_WR_means, main = "Accelerometer Right Wrist", xlab = "", ylab = "Mean Value", cex.main = 1.5, xaxt = "n")
lines(Acc_WR_means, type = "o", col = "#1f77b4", lwd = 2)
points(highlight_points, Acc_WR_means[highlight_points], col = "#ff7f7f", pch = 16, cex = 1.5)
abline(v = 23.5, lty = 2, col = "red", lwd = 2)
axis(1, at = custom_positions, labels = custom_labels)

# Performance Plot
plot(Perf_means, main = "Gaming Score", xlab = "", ylab = "Mean Value", cex.main = 1.5, xaxt = "n")
lines(Perf_means, type = "o", col = "#1f77b4", lwd = 2)
custom_labels3 <- c(1, 2, 3, 1, 2, 3, 4, 5, 6, 7, 8)
axis(1, at = 1:length(custom_labels3), labels = custom_labels3)
abline(v = 3.5, lty = 2, col = "red")


# Affect Plot

par(mar = c(5, 4, 4, 8) + 0.1)
plot(Amus_means, type = "o", col = "purple",lwd = 2, ylim = range(c(Amus_means, Joy_means, Exc_means, Pride_means, Stress_means, Anger_means, Fear_means, Over_means)),
     main = "Affect", xlab = "", ylab = "Mean Value", xaxt = "n", cex.main = 1.5)
custom_labels2 <- c(1, 2, 3, 1, 2, 3, 4, 5, 6, 7, 8)
axis(1, at = 1:length(custom_labels2), labels = custom_labels2)

lines(Joy_means, type = "o", col = "blue",lwd = 2)
lines(Exc_means, type = "o", col = "green",lwd = 2)
lines(Pride_means, type = "o", col = "yellow",lwd = 2)
lines(Stress_means, type = "o", col = "black",lwd = 2)
lines(Anger_means, type = "o", col = "pink",lwd = 2)
lines(Fear_means, type = "o", col = "gray",lwd = 2)
lines(Over_means, type = "o", col = "red", lwd = 2)

# Add a vertical line
abline(v = 3.5, lty = 2, col = "blue")

# Add the legend closer to the main graph and aligning its height
legend("topright", inset = c(-0.3, 0), legend = c("Amusment", "Joy", "Excitement", "Pride", "Stress", "Anger", "Fear", "Overwhelm"),
       col = c("purple", "blue", "green", "yellow", "black", "pink", "gray", "red"), lty = 1, xpd = TRUE, bty = "n")



```
